---
title: "ind.sat.data.processing"
author: "Aarsh Batra"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load_libraries}

#> libraries
library(tidyverse)
library(jsonlite)
library(here)
library(sf)
library(ggplot2)
library(geojson)
library(devtools)
library(geojsonio)
library(geojsonsf)
library(roxygen2)
library(magrittr)
library(stringr)
library(data.table)
library(viridis)
library(ggthemes)
library(hrbrthemes)
library(RColorBrewer)
library(haven)
library(janitor)
library(lwgeom)
library(rvest)
library(dplyr)
library(ggrepel)
library(ggsflabel)
library(fuzzyjoin)
library(tidytext)
library(tm)
library(wordcloud2)
library(recipes)
library(forcats)
library(sjlabelled) 
library(geoarrow) # for exporting shapefiles to a geoparquet format
library(paletteer)
library(jpeg)
library(patchwork)
library(ggtext)
library(readxl)
library(tictoc)
library(googlesheets4)
library(reticulate)
library(raster)
library(fasterize)
library(microbenchmark)
library(future)
library(raster)
library(rasterVis)
library(ncdf4)
library(lattice)
library(foster)
library(rgdal)

# load and automatically document all functions
load_all()
document()


#> Negation of the "%in%" operator function
`%notin%` <- Negate(`%in%`)

#> shrug district level shapefile with uid for rasterization
shrug_master_district_shp <- st_read(str_c(here(), "/data/raw/shrug/shrug_master_district_shp_with_uid_for_rast_wa.gpkg"))

#> shrug subdistrict level shapefile with uid for rasterization
shrug_master_subdist_shp <- st_read(str_c(here(), "/data/raw/shrug/shrug_master_subdistrict_shp_with_uid_for_rast_wa.gpkg"))

shrug_master_subdist_shp <- shrug_master_subdist_shp %>%
  dplyr::mutate(uid_for_rasterization = row_number()) %>%
  dplyr::select(-geom, geom) %>%
  st_as_sf()

#> state level shapefile
ind_st_level_shp <- st_read(str_c(here(), "/data/raw/other.datasets/shpfiles/global_state.shp"))

ind_st_level_shp <- ind_st_level_shp %>%
  dplyr::filter(name0 == "India")
  

```

## Housekeeping items: set reference datasets that will be used throughout in the pipeline ahead
```{r housekeeping}


#> [To be set by user]set the ref admin layer to be used for cropping and masking global rasters: using less detailed layers make cropping and masking more efficient later on
ref_admin_level_shp_file <- shrug_master_district_shp

#> [To be set by user] set the ref admin layer
ref_shp_file_to_be_rasterized <- shrug_master_subdist_shp


```

## Process satellite derived PM2.5 data and produced population weighted PM2.5 at India block level
```{r process_satellite_data_in_loop}


#> load in a single pollution raster, and crop it to India

pol_raster_path_vec <- list.files("./data/raw/annual_pm2.5_v6.gl.02_raw_rasters_0.01_0.01/annual", full.names = TRUE) 

yearly_data_processed_list <- list()

for(i in 1 : length(pol_raster_path_vec)){

tic()
  
  cur_rast_year <- as.numeric(str_remove(str_extract(pol_raster_path_vec[i], "\\d+\\.nc"), "12.nc"))

  print(str_c(cur_rast_year, " (", i, "/", length(pol_raster_path_vec), ") pm2.5 pol data processing r1 starts!"))
    
  unit_pol_raster <- map(pol_raster_path_vec[i], ~nc_to_raster_layer(., "PM25", "latitude", "longitude"), .progress = TRUE)

#> crop the unit_raster using ref shp
unit_pol_raster_cropped_admin_lev_shp_file <- raster::crop(unit_pol_raster[[1]], ref_admin_level_shp_file)

#> mask the unit raster using ref shp
unit_pol_raster_cropped_admin_lev_shp_file_mask <- mask(unit_pol_raster_cropped_admin_lev_shp_file, ref_admin_level_shp_file)

#> load in the population raster
unit_pop_raster <- raster("./data/raw/landscan-global-2022-assets/landscan-global-2022.tif")

#> crop the population raster
unit_pop_raster_cropped_admin_lev_shp_file <- raster::crop(unit_pop_raster, ref_admin_level_shp_file)

#> mask the population raster
unit_pop_raster_cropped_admin_lev_shp_file_mask <- mask(unit_pop_raster_cropped_admin_lev_shp_file, ref_admin_level_shp_file)

#> match pol raster to pop raster
unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop <- matchResolution(unit_pol_raster_cropped_admin_lev_shp_file_mask, unit_pop_raster_cropped_admin_lev_shp_file_mask)

#> fasterize the village level shapefile (with a uid for rasterization) to the same resolution as the pop, pol rasters
sh_admin_int_shp_rasterized <- fasterize(ref_shp_file_to_be_rasterized, unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop, field = "uid_for_rasterization", fun = "last")

#> set names for rasters

names(unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop) <- "pol_rast"
names(unit_pop_raster_cropped_admin_lev_shp_file_mask) <- "pop_rast"
names(sh_admin_int_shp_rasterized) <- "sh_rast"

#> create a raster brick out of pol rast, pop rast, sh vil rast
region_raster_brick <- sh_admin_int_shp_rasterized %>%
  addLayer(unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop) %>%
  addLayer(unit_pop_raster_cropped_admin_lev_shp_file_mask)

#> create a df out of the brick
region_raster_brick_df <- raster::as.data.frame(region_raster_brick)
region_raster_brick_df <- region_raster_brick_df %>%
  dplyr::filter(!is.na(pol_rast), !is.na(pop_rast))

#> create an arrow table from the df before collapsing
region_raster_brick_df_arrow <- region_raster_brick_df %>%
  arrow::as_arrow_table()

#> collapse arrow table
region_raster_brick_df_arrow_collapse <- region_raster_brick_df_arrow %>%
  # dplyr::filter((!is.na(sh_rast)) & ((as.character(sh_rast) != "NA"))) %>%
  dplyr::group_by(sh_rast) %>%
  dplyr::collect() %>%
  dplyr::mutate(pop_weights = pop_rast/sum(pop_rast, na.rm = TRUE), 
         pollution_pop_weighted = pop_weights*pol_rast) %>%
  dplyr::summarise(total_population = sum(pop_rast, na.rm = TRUE), 
            avg_pm2.5_pollution = sum(pollution_pop_weighted, na.rm = TRUE)) %>%
  dplyr::ungroup() 

#> shrug ids not captured in the first round of processing
uids_not_captured_r1 <- ref_shp_file_to_be_rasterized$uid_for_rasterization[ref_shp_file_to_be_rasterized$uid_for_rasterization %notin% region_raster_brick_df_arrow_collapse$sh_rast]

#> uids with NA pop
uids_with_na_pop <- region_raster_brick_df_arrow_collapse %>% dplyr::filter(is.na(total_population)) %>% 
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> uids with NA pol
uids_with_na_pol <- region_raster_brick_df_arrow_collapse %>% dplyr::filter(is.na(avg_pm2.5_pollution)) %>% 
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> uids with 0 population
uids_with_zero_pop <- region_raster_brick_df_arrow_collapse %>%
  dplyr::filter(total_population == 0) %>%
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> concat all unprocessed shrud ids into a single vector
uids_not_captured_master_r1 <- c(uids_not_captured_r1, uids_with_na_pop, uids_with_na_pol, uids_with_zero_pop)
uids_not_captured_master_r1 <- uids_not_captured_master_r1 %>%
  unique()


### Process unprocess shrug ids by resampling======================================================

#> resampling details

print(str_c(cur_rast_year, "capturing residual tiny blocks..."))

resample_to_res <- 0.001 
res_resample_from <- 0.00833333
resample_obj_capture_list <- list()

#> restricting the shp file to the unprocessed shr ids that will be processed one at a time
sh_admin_int_shp_unprocessed  <- ref_shp_file_to_be_rasterized  %>%
  filter(uid_for_rasterization %in% uids_not_captured_master_r1)



#> reprocessing unprocessed shr ids one at a time
for (j in 1 : length(uids_not_captured_master_r1)){
  
  # rasterize shpfile to the current unprocessed shr id
  sh_admin_int_shp_unprocessed_rasterized <- fasterize(sh_admin_int_shp_unprocessed %>% filter(uid_for_rasterization == uids_not_captured_master_r1[j]), raster::raster(ext = raster::extent(sh_admin_int_shp_unprocessed %>% filter(uid_for_rasterization == uids_not_captured_master_r1[j])), resolution = resample_to_res, crs = crs(sh_admin_int_shp_unprocessed)), field = "uid_for_rasterization", fun = "last")
  
  # set name of the rasterized shapefile
  names(sh_admin_int_shp_unprocessed_rasterized) <- "sh_rast"
  
  # crop pol raster to the cur unprocessed shr id
  unit_pol_raster_crop_unproc <- crop(unit_pol_raster[[1]], sh_admin_int_shp_unprocessed_rasterized)
  
  # resample to the new resolution
  unit_pol_raster_crop_unproc_resample <- raster::resample(unit_pol_raster_crop_unproc, sh_admin_int_shp_unprocessed_rasterized, method = "ngb")
  
  # mask resampled pol raster to the cur unprocessed shr id
   unit_pol_raster_crop_unproc_resample_mask <- raster::mask(unit_pol_raster_crop_unproc_resample, sh_admin_int_shp_unprocessed_rasterized)
   
   
    # crop pop raster to the cur unprocessed shr id
  unit_pop_raster_crop_unproc <- crop(unit_pop_raster, sh_admin_int_shp_unprocessed_rasterized)
  
  # replace population values with population densities before resampling
  raster::values(unit_pop_raster_crop_unproc) <- as.vector(unit_pop_raster_crop_unproc * (((resample_to_res)^2)/((res_resample_from)^2)))
  
  # resample to the new resolution
  unit_pop_raster_crop_unproc_resample <- raster::resample(unit_pop_raster_crop_unproc, sh_admin_int_shp_unprocessed_rasterized, method = "ngb")
  
  # mask resampled pol raster to the cur unprocessed shr id
   unit_pop_raster_crop_unproc_resample_mask <- raster::mask(unit_pop_raster_crop_unproc_resample, sh_admin_int_shp_unprocessed_rasterized)
   
   
   #> set names for rasters

names(unit_pol_raster_crop_unproc_resample_mask) <- "pol_rast"
names(unit_pop_raster_crop_unproc_resample_mask) <- "pop_rast"
names(sh_admin_int_shp_unprocessed_rasterized) <- "sh_rast"
   
  
  #> create a raster brick out of pol rast, pop rast, sh vil rast
region_raster_brick_resample <- sh_admin_int_shp_unprocessed_rasterized %>%
  addLayer(unit_pol_raster_crop_unproc_resample_mask) %>%
  addLayer(unit_pop_raster_crop_unproc_resample_mask)

#> create a df out of the brick
region_raster_brick_df_resample <- raster::as.data.frame(region_raster_brick_resample)
region_raster_brick_df_resample <- region_raster_brick_df_resample %>%
  dplyr::filter(!is.na(pol_rast), !is.na(pop_rast))

#> create an arrow table from the df before collapsing
region_raster_brick_df_arrow_resample <- region_raster_brick_df_resample %>%
  arrow::as_arrow_table()

#> collapse arrow table
region_raster_brick_df_arrow_collapse_resample <- region_raster_brick_df_arrow_resample %>%
  # dplyr::filter((!is.na(sh_rast)) & ((as.character(sh_rast) != "NA"))) %>%
  dplyr::group_by(sh_rast) %>%
  dplyr::collect() %>%
  dplyr::mutate(pop_weights = pop_rast/sum(pop_rast, na.rm = TRUE), 
         pollution_pop_weighted = pop_weights*pol_rast) %>%
  dplyr::summarise(total_population = sum(pop_rast, na.rm = TRUE), 
            avg_pm2.5_pollution = sum(pollution_pop_weighted, na.rm = TRUE)) %>%
  dplyr::ungroup() 

resample_obj_capture_list[[j]] <- region_raster_brick_df_arrow_collapse_resample

print(str_c(j, " out of ", length(uids_not_captured_master_r1), " tiny blocks reprocessed!"))


}

#> rbind all residual blocks captured in r2 of processing
resample_obj_capture_df <- resample_obj_capture_list %>%
  rbindlist() %>%
  as_tibble()

#> rbind the residual blocks processed in r2 with the blocks from r1
region_raster_brick_df_arrow_collapse_final <- region_raster_brick_df_arrow_collapse %>%
  rbind(resample_obj_capture_df) %>%
  arrange(sh_rast)
  
#> rename pollution column
colnames(region_raster_brick_df_arrow_collapse_final)[colnames(region_raster_brick_df_arrow_collapse_final) == "avg_pm2.5_pollution"] <- str_c("avg_pm2.5_", cur_rast_year)

yearly_data_processed_list[[i]] <- region_raster_brick_df_arrow_collapse_final

print(str_c(cur_rast_year, " pm2.5 pol data processing r1 completed!"))




toc()  
}



```

## create a master final SHRUG mapped dataset

```{r master_sh_map_data_create}

# creating a master dataset for all years from 1990 to 2020
final_all_years_cbind_df <- reduce(yearly_data_processed_list, function(x, y) left_join(x, y, by = c("sh_rast", "total_population")))

# adding SHRUG admin area names from the block level shapefile
final_all_years_cbind_df_sh_mapped <- final_all_years_cbind_df %>%
  left_join(shrug_master_subdist_shp %>% st_drop_geometry(), by = c("sh_rast" = "uid_for_rasterization")) %>%
  dplyr::select(pc11_state_id, pc11_district_id, pc11_subdistrict_id, state_name, district_name, subdistrict_name, total_population, everything()) %>%
  dplyr::select(-c(composite_uid, sh_rast)) %>%
  rename(subdistrict_population = total_population)

```

## [EXTRA CODE FOR PLOTTING, not part of main pipeline] Plotting a sample map: annual avg pm2.5 in 2022 in India at block level in micrograms per cubic meter

```{r eval=FALSE, include=FALSE}

# plt_pol_1998 <- final_all_years_cbind_df_sh_mapped  %>% 
#   left_join(shrug_master_subdist_shp %>% 
#               dplyr::select(state_name, district_name, subdistrict_name, geom), by = c("state_name", "district_name", "subdistrict_name"))  %>% 
#   st_as_sf() %>% ggplot() + geom_sf(mapping = aes(fill = avg_pm2.5_1998), color = "transparent") + scale_fill_viridis(option = "magma", direction = -1) +
#   geom_sf(data = ind_st_level_shp, color = "black", lwd = 0.01, fill = "transparent") +
#   theme_map() +
#   labs(fill = expression("Annual average PM"[2.5] ~ "(µg/" * m^3 * ")")) +
#  theme(legend.position = "bottom",           
#         legend.direction = "horizontal", 
#        legend.key.width = unit(3, "cm"), 
#        plot.background = element_rect(fill = "white", color = "white")) +
#      guides(fill = guide_colorbar(title.position = "top"))
# 
# plt_pol_2010 <- final_all_years_cbind_df_sh_mapped  %>% 
#   left_join(shrug_master_subdist_shp %>% 
#               dplyr::select(state_name, district_name, subdistrict_name, geom), by = c("state_name", "district_name", "subdistrict_name"))  %>% 
#   st_as_sf() %>% ggplot() + geom_sf(mapping = aes(fill = avg_pm2.5_2010), color = "transparent") + scale_fill_viridis(option = "magma", direction = -1) +
#   geom_sf(data = ind_st_level_shp, color = "black", lwd = 0.01, fill = "transparent") +
#   theme_map() +
#   labs(fill = expression("Annual average PM"[2.5] ~ "(µg/" * m^3 * ")")) +
#  theme(legend.position = "bottom",           # Place legend at the bottom
#         legend.direction = "horizontal", 
#        legend.key.width = unit(3, "cm"), 
#        plot.background = element_rect(fill = "white", color = "white")) +
#     guides(fill = guide_colorbar(title.position = "top"))     
#      
#      
# plt_pol_2022 <- final_all_years_cbind_df_sh_mapped  %>% 
#   left_join(shrug_master_subdist_shp %>% 
#               dplyr::select(state_name, district_name, subdistrict_name, geom), by = c("state_name", "district_name", "subdistrict_name"))  %>% 
#   st_as_sf() %>% ggplot() + geom_sf(mapping = aes(fill = avg_pm2.5_2022), color = "transparent") + scale_fill_viridis(option = "magma", direction = -1, breaks = c(0, 5, 10, 20, 40, 60, 80, 120), limits = c(0, 120)) +
#   geom_sf(data = ind_st_level_shp, color = "black", lwd = 0.12, fill = "transparent") +
#   theme_map() +
#   labs(fill = expression("Annual Average"~PM[2.5]~"(micrograms per cubic metre)"), 
#        title = expression("Block level"~PM[2.5]~"pollution in India, 2022")) +
#  theme(legend.position = "bottom",           # Place legend at the bottom
#         legend.direction = "horizontal", 
#        legend.key.width = unit(3, "cm"), 
#        plot.background = element_rect(fill = "white", color = "white"), 
#        legend.title = element_text(size = 14, family = "Mono"), 
#        legend.text = element_text(size = 11, family = "Mono"), 
#        plot.title = element_text(size = 18, family = "Mono", hjust = 0)) +
#     guides(fill = guide_colorbar(title.position = "top"))
# 
# plt_final <- ggarrange(plt_pol_1998, plt_pol_2022, common.legend = TRUE, labels = c("1998", "2022"), legend = "top")

```


