---
title: "22.ind.block.pm2.5.sat.data.processing.1998.2024"
author: "Aarsh Batra"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load_libraries}

#> libraries
library(tidyverse)
library(jsonlite)
library(here)
library(sf)
library(ggplot2)
library(geojson)
library(devtools)
library(geojsonio)
library(geojsonsf)
library(roxygen2)
library(magrittr)
library(stringr)
library(data.table)
library(viridis)
library(ggthemes)
library(hrbrthemes)
library(RColorBrewer)
library(haven)
library(janitor)
library(lwgeom)
library(rvest)
library(dplyr)
library(ggrepel)
library(ggsflabel)
library(fuzzyjoin)
library(tidytext)
library(tm)
library(wordcloud2)
library(recipes)
library(forcats)
library(sjlabelled) 
library(geoarrow) # for exporting shapefiles to a geoparquet format
library(paletteer)
library(jpeg)
library(patchwork)
library(ggtext)
library(readxl)
library(tictoc)
library(googlesheets4)
library(reticulate)
library(raster)
library(fasterize)
library(microbenchmark)
library(future)
library(raster)
library(rasterVis)
library(ncdf4)
library(lattice)
library(foster)
library(rgdal)
```


## Load relevant functions, global variables and parameters setup
```{r}
# load and automatically document all functions
load_all()
document()


#> Negation of the "%in%" operator function
`%notin%` <- Negate(`%in%`)

#> shrug district level shapefile with uid for rasterization
shrug_master_district_shp <- st_read(str_c(here(), "/data/raw/shrug/shrug_master_district_shp_with_uid_for_rast_wa.gpkg"))

#> shrug subdistrict level shapefile with uid for rasterization
shrug_master_subdist_shp <- st_read(str_c(here(), "/data/raw/shrug/shrug_master_subdistrict_shp_with_uid_for_rast_wa.gpkg"))

shrug_master_subdist_shp <- shrug_master_subdist_shp %>%
  dplyr::mutate(uid_for_rasterization = row_number()) %>%
  dplyr::select(-geom, geom) %>%
  st_as_sf()

#> state level shapefile
ind_st_level_shp <- st_read(str_c(here(), "/data/raw/other.datasets/shpfiles/global_state.shp"))

ind_st_level_shp <- ind_st_level_shp %>%
  dplyr::filter(name0 == "India")


#> load in a single pollution raster, and crop it to India
#> 
pol_raster_path_vec <- list.files("./data/raw/monthly_pm2.5_v5.gl.06_raw_rasters_0.01_0.01_1998_2024_asia/Monthly/", full.names = TRUE, recursive = TRUE) 

#> load in the population raster
unit_pop_raster <- raster("./data/raw/landscan-global-2024-assets/landscan-global-2024.tif")

yearly_data_processed_list <- list()
  
#> [To be set by user]set the ref admin layer to be used for cropping and masking global rasters: using less detailed layers make cropping and masking more efficient later on
ref_admin_level_shp_file <- shrug_master_district_shp

#> [To be set by user] set the ref admin layer
ref_shp_file_to_be_rasterized <- shrug_master_subdist_shp


#> Set parameters for raw pollution rasters (check for a given raster once a year once data is downloaded from ACAG website)
pm25_rast_name <- "GWRPM25"
lat_rast_name <- "lat"
lon_rast_name <- "lon"

```


## Process satellite derived PM2.5 data and produced population weighted PM2.5 at Pakistan tehsil level by year in a loop
```{r process_satellite_data_in_loop}

for(i in 143 : length(pol_raster_path_vec)){

## testing
# if(i == 7){
#   break
# }  
  
tic()
  
  cur_rast_year <- as.numeric(str_remove(str_extract(pol_raster_path_vec[i], "\\d+\\.nc"), "\\d{2}.nc"))
  
  cur_rast_month <- as.numeric(str_remove((str_extract(str_extract(pol_raster_path_vec[i], "\\d+\\.nc"), "\\d{2}.nc")), ".nc"))
  
  cur_rast_month <- str_to_lower(month.name[cur_rast_month])

  print(str_c(cur_rast_year, "-", cur_rast_month, ": ", " (", i, "/", length(pol_raster_path_vec), ") pm2.5 pol data processing r1 starts!"))
    
  unit_pol_raster <- map(pol_raster_path_vec[i], ~nc_to_raster_layer(., pm25_rast_name, lat_rast_name, lon_rast_name), .progress = TRUE)

#> crop the unit_raster using ref shp
unit_pol_raster_cropped_admin_lev_shp_file <- raster::crop(unit_pol_raster[[1]], ref_admin_level_shp_file)

#> mask the unit raster using ref shp
unit_pol_raster_cropped_admin_lev_shp_file_mask <- mask(unit_pol_raster_cropped_admin_lev_shp_file, ref_admin_level_shp_file)

#> clamp the pol raster, to replace any negative value pixels with pixels of value = 0, as pollution concentrations cannot have negative values
unit_pol_raster_cropped_admin_lev_shp_file_mask <- terra::clamp(
  unit_pol_raster_cropped_admin_lev_shp_file_mask,
  lower = 0,  # Any value below 0 will be set to 0
  values = TRUE  # Apply the threshold and replace values below 0 with 0
)

#> crop the population raster
unit_pop_raster_cropped_admin_lev_shp_file <- raster::crop(unit_pop_raster, ref_admin_level_shp_file)

#> mask the population raster
unit_pop_raster_cropped_admin_lev_shp_file_mask <- mask(unit_pop_raster_cropped_admin_lev_shp_file, ref_admin_level_shp_file)

#> clamp the pop raster, to replace any negative value pixels with pixels of value = 0, as population cannot have negative values
unit_pop_raster_cropped_admin_lev_shp_file_mask <- terra::clamp(
 unit_pop_raster_cropped_admin_lev_shp_file_mask,
  lower = 0,  # Any value below 0 will be set to 0
  values = TRUE  # Apply the threshold and replace values below 0 with 0
)


#> match pol raster to pop raster
unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop <- matchResolution(unit_pol_raster_cropped_admin_lev_shp_file_mask, unit_pop_raster_cropped_admin_lev_shp_file_mask)

#> fasterize the village level shapefile (with a uid for rasterization) to the same resolution as the pop, pol rasters
sh_admin_int_shp_rasterized <- fasterize(ref_shp_file_to_be_rasterized, unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop, field = "uid_for_rasterization", fun = "last")

#> set names for rasters

names(unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop) <- "pol_rast"
names(unit_pop_raster_cropped_admin_lev_shp_file_mask) <- "pop_rast"
names(sh_admin_int_shp_rasterized) <- "sh_rast"

#> create a raster brick out of pol rast, pop rast, sh vil rast
region_raster_brick <- sh_admin_int_shp_rasterized %>%
  addLayer(unit_pol_raster_cropped_admin_lev_shp_file_mask_match_pop) %>%
  addLayer(unit_pop_raster_cropped_admin_lev_shp_file_mask)

#> create a df out of the brick
region_raster_brick_df <- raster::as.data.frame(region_raster_brick)
region_raster_brick_df <- region_raster_brick_df %>%
  dplyr::filter(!is.na(pol_rast), !is.na(pop_rast))

#> create an arrow table from the df before collapsing
region_raster_brick_df_arrow <- region_raster_brick_df %>%
  arrow::as_arrow_table()

#> collapse arrow table
region_raster_brick_df_arrow_collapse <- region_raster_brick_df_arrow %>%
  # dplyr::filter((!is.na(sh_rast)) & ((as.character(sh_rast) != "NA"))) %>%
  dplyr::group_by(sh_rast) %>%
  dplyr::collect() %>%
  dplyr::mutate(pop_weights = pop_rast/sum(pop_rast, na.rm = TRUE), 
         pollution_pop_weighted = pop_weights*pol_rast) %>%
  dplyr::summarise(total_population = sum(pop_rast, na.rm = TRUE), 
            avg_pm2.5_pollution = sum(pollution_pop_weighted, na.rm = TRUE)) %>%
  dplyr::ungroup() 

#> shrug ids not captured in the first round of processing
uids_not_captured_r1 <- ref_shp_file_to_be_rasterized$uid_for_rasterization[ref_shp_file_to_be_rasterized$uid_for_rasterization %notin% region_raster_brick_df_arrow_collapse$sh_rast]

#> uids with NA pop
uids_with_na_pop <- region_raster_brick_df_arrow_collapse %>% dplyr::filter(is.na(total_population)) %>% 
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> uids with NA pol
uids_with_na_pol <- region_raster_brick_df_arrow_collapse %>% dplyr::filter(is.na(avg_pm2.5_pollution)) %>% 
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> uids with 0 population
uids_with_zero_pop <- region_raster_brick_df_arrow_collapse %>%
  dplyr::filter(total_population == 0) %>%
  dplyr::select(sh_rast) %>%
  unlist() %>%
  as.vector()

#> concat all unprocessed shrud ids into a single vector
uids_not_captured_master_r1 <- c(uids_not_captured_r1, uids_with_na_pop, uids_with_na_pol, uids_with_zero_pop)
uids_not_captured_master_r1 <- uids_not_captured_master_r1 %>%
  unique()

#> resample and capture uncaptured blocks
if(length(uids_not_captured_master_r1) == 0){
  print("all blocks captured in first round, resampling not needed")
  region_raster_brick_df_arrow_collapse_final <- region_raster_brick_df_arrow_collapse

#> rename pollution column
colnames(region_raster_brick_df_arrow_collapse_final)[colnames(region_raster_brick_df_arrow_collapse_final) == "avg_pm2.5_pollution"] <- str_c("avg_pm2.5_", cur_rast_year)

yearly_data_processed_list[[i]] <- region_raster_brick_df_arrow_collapse_final

print(str_c(cur_rast_year, " pm2.5 pol data processing r1 completed!"))


  next
} else {
  


### Process unprocess shrug ids by resampling======================================================

#> resampling details

print(str_c(cur_rast_year, "capturing residual tiny blocks..."))

resample_to_res <- 0.001 
res_resample_from <- 0.00833333
resample_obj_capture_list <- list()

#> restricting the shp file to the unprocessed shr ids that will be processed one at a time
sh_admin_int_shp_unprocessed  <- ref_shp_file_to_be_rasterized  %>%
  filter(uid_for_rasterization %in% uids_not_captured_master_r1)


if(length(sh_admin_int_shp_unprocessed))

#> reprocessing unprocessed shr ids one at a time
for (j in 1 : length(uids_not_captured_master_r1)){
  
  # rasterize shpfile to the current unprocessed shr id
  sh_admin_int_shp_unprocessed_rasterized <- fasterize(sh_admin_int_shp_unprocessed %>% filter(uid_for_rasterization == uids_not_captured_master_r1[j]), raster::raster(ext = raster::extent(sh_admin_int_shp_unprocessed %>% filter(uid_for_rasterization == uids_not_captured_master_r1[j])), resolution = resample_to_res, crs = crs(sh_admin_int_shp_unprocessed)), field = "uid_for_rasterization", fun = "last")
  
  # set name of the rasterized shapefile
  names(sh_admin_int_shp_unprocessed_rasterized) <- "sh_rast"
  
  # crop pol raster to the cur unprocessed shr id
  unit_pol_raster_crop_unproc <- raster::crop(unit_pol_raster[[1]], sh_admin_int_shp_unprocessed_rasterized)
  
  # resample to the new resolution
  unit_pol_raster_crop_unproc_resample <- raster::resample(unit_pol_raster_crop_unproc, sh_admin_int_shp_unprocessed_rasterized, method = "ngb")
  
  # mask resampled pol raster to the cur unprocessed shr id
   unit_pol_raster_crop_unproc_resample_mask <- raster::mask(unit_pol_raster_crop_unproc_resample, sh_admin_int_shp_unprocessed_rasterized)
   
  #> clamp the pol raster, to replace any negative value pixels with pixels of value = 0, as pollution concentrations cannot have negative values
unit_pol_raster_crop_unproc_resample_mask <- terra::clamp(
  unit_pol_raster_crop_unproc_resample_mask,
  lower = 0,  # Any value below 0 will be set to 0
  values = TRUE  # Apply the threshold and replace values below 0 with 0
) 
   
    # crop pop raster to the cur unprocessed shr id
  unit_pop_raster_crop_unproc <- crop(unit_pop_raster, sh_admin_int_shp_unprocessed_rasterized)
  
  # replace population values with population densities before resampling
  raster::values(unit_pop_raster_crop_unproc) <- as.vector(unit_pop_raster_crop_unproc * (((resample_to_res)^2)/((res_resample_from)^2)))
  
  # resample to the new resolution
  unit_pop_raster_crop_unproc_resample <- raster::resample(unit_pop_raster_crop_unproc, sh_admin_int_shp_unprocessed_rasterized, method = "ngb")
  
  # mask resampled pop raster to the cur unprocessed shr id
   unit_pop_raster_crop_unproc_resample_mask <- raster::mask(unit_pop_raster_crop_unproc_resample, sh_admin_int_shp_unprocessed_rasterized)
   
    #> clamp the pop raster, to replace any negative value pixels with pixels of value = 0, as population cannot have negative values
unit_pop_raster_crop_unproc_resample_mask <- terra::clamp(
  unit_pop_raster_crop_unproc_resample_mask,
  lower = 0,  # Any value below 0 will be set to 0
  values = TRUE  # Apply the threshold and replace values below 0 with 0
) 
   #> set names for rasters

names(unit_pol_raster_crop_unproc_resample_mask) <- "pol_rast"
names(unit_pop_raster_crop_unproc_resample_mask) <- "pop_rast"
names(sh_admin_int_shp_unprocessed_rasterized) <- "sh_rast"
   
  
  #> create a raster brick out of pol rast, pop rast, sh vil rast
region_raster_brick_resample <- sh_admin_int_shp_unprocessed_rasterized %>%
  addLayer(unit_pol_raster_crop_unproc_resample_mask) %>%
  addLayer(unit_pop_raster_crop_unproc_resample_mask)

#> create a df out of the brick
region_raster_brick_df_resample <- raster::as.data.frame(region_raster_brick_resample)
region_raster_brick_df_resample <- region_raster_brick_df_resample %>%
  dplyr::filter(!is.na(pol_rast), !is.na(pop_rast))

#> create an arrow table from the df before collapsing
region_raster_brick_df_arrow_resample <- region_raster_brick_df_resample %>%
  arrow::as_arrow_table()

#> collapse arrow table
region_raster_brick_df_arrow_collapse_resample <- region_raster_brick_df_arrow_resample %>%
  # dplyr::filter((!is.na(sh_rast)) & ((as.character(sh_rast) != "NA"))) %>%
  dplyr::group_by(sh_rast) %>%
  dplyr::collect() %>%
  dplyr::mutate(pop_weights = pop_rast/sum(pop_rast, na.rm = TRUE), 
         pollution_pop_weighted = pop_weights*pol_rast) %>%
  dplyr::summarise(total_population = sum(pop_rast, na.rm = TRUE), 
            avg_pm2.5_pollution = sum(pollution_pop_weighted, na.rm = TRUE)) %>%
  dplyr::ungroup() 

resample_obj_capture_list[[j]] <- region_raster_brick_df_arrow_collapse_resample

print(str_c(j, " out of ", length(uids_not_captured_master_r1), " tiny blocks reprocessed!"))


}

#> rbind all residual blocks captured in r2 of processing
resample_obj_capture_df <- resample_obj_capture_list %>%
  rbindlist() %>%
  as_tibble()

#> rbind the residual blocks processed in r2 with the blocks from r1
region_raster_brick_df_arrow_collapse_final <- region_raster_brick_df_arrow_collapse %>%
  rbind(resample_obj_capture_df) %>%
  arrange(sh_rast)

}

#> rename pollution column
colnames(region_raster_brick_df_arrow_collapse_final)[colnames(region_raster_brick_df_arrow_collapse_final) == "avg_pm2.5_pollution"] <- str_c("avg_pm2.5_", cur_rast_year, "_", cur_rast_month)

yearly_data_processed_list[[i]] <- region_raster_brick_df_arrow_collapse_final

yearly_data_processed_list[[i]] %>%
  writexl::write_xlsx(paste0("./data/intermediate/22.ind.block.pm2.5.sat.data.processing.1998.2024/monthly_pm2.5_v5.gl.06_0.01_0.01_1998_2024_asia/year-month-datasets/", "avg_pm2.5_", cur_rast_year, "_", cur_rast_month, ".xlsx"))

print(str_c(cur_rast_year, "-", cur_rast_month, ": ", " pm2.5 pol data processing fully completed and files preserved!"))




toc()  
}



```

## create a master final SHRUG mapped dataset

```{r master_sh_map_data_create}

#### uncomment and run this (after updating appropriate ver of dataset) if you have already processed the month-year files above and want to directly load and perform remaining steps to finalize. 
# yearly_data_processed_list <- map(list.files("./data/intermediate/22.ind.block.pm2.5.sat.data.processing.1998.2024/monthly_pm2.5_v5.gl.06_0.01_0.01_1998_2024_asia/year-month-datasets/", full.names = TRUE), ~readxl::read_excel(.x), .progress = TRUE)

# removing any duplicate regions
yearly_data_processed_list_fin <- yearly_data_processed_list %>% map(~.x %>% distinct_all())

# adding a NA for pol where while population is 0
yearly_data_processed_list_fin <- yearly_data_processed_list_fin %>%   map(~ .x %>%
        mutate(across(starts_with("avg_pm2.5"),
                      ~ na_if(., 0))))

# creating a master dataset for all years from 1998 to 2024
final_all_years_cbind_df <- reduce(yearly_data_processed_list_fin, function(x, y) left_join(x, y, by = c("sh_rast", "total_population")))

# adding SHRUG admin area names from the block level shapefile
final_all_years_cbind_df_sh_mapped <- final_all_years_cbind_df %>%
  left_join(ref_shp_file_to_be_rasterized %>% st_drop_geometry(), by = c("sh_rast" = "uid_for_rasterization")) %>%
  dplyr::select(state_name, district_name, subdistrict_name, total_population, starts_with("avg_pm")) %>%
  rename(subdistrict_population = total_population)

final_all_years_cbind_df %>%  write_csv(paste0("./data/processed/22.ind.block.pm2.5.sat.data.processing.1998.2024/av_pm2.5_ind_blk_mth_1998_2024_wide.csv"))

# long form version of the dataset
final_all_years_cbind_df_sh_mapped_long <- final_all_years_cbind_df_sh_mapped %>% pivot_longer(cols = !c(state_name, district_name, subdistrict_name, subdistrict_population), names_to = c("year", "month"), names_pattern = "(\\d{4})_(.+)", values_to = "avg_pm2.5") %>%
  dplyr::mutate(year = as.numeric(year)) %>%
     dplyr::mutate(month = factor(month, 
                          levels = c("january", "february", "march", "april", "may", "june", 
                                     "july", "august", "september", "october", "november", "december"))) 

final_all_years_cbind_df_sh_mapped_long %>%  write_csv(paste0("./data/processed/22.ind.block.pm2.5.sat.data.processing.1998.2024/av_pm2.5_ind_blk_mth_1998_2024_long.csv"))

```


## [EXTRA CODE FOR PLOTTING, not part of main pipeline] Plotting a sample map: monthly avg pm2.5 in all years in a loop in India at block level in micrograms per cubic meter

```{r}

#### plot monthly screenshot for all years for India at block level----------------------
ref_yr <- c(1998:2024)


for(i in 1 : length(ref_yr)){
plt1 <- final_all_years_cbind_df_sh_mapped_long %>%
  mutate(month = factor(month, levels = str_to_lower(month.name), labels = month.name)) %>%
  filter(year %in% c(ref_yr[i])) %>%
  left_join(
    shrug_master_subdist_shp %>%
      dplyr::select(state_name, district_name, subdistrict_name, geom), 
    by = c("state_name", "district_name", "subdistrict_name")
  ) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = avg_pm2.5), color = NA) +
  scale_fill_stepsn(
    colors = rev(viridisLite::magma(8)),
    breaks = c(0, 20, 40, 60, 100, 140, 220, 300),
    limits = c(0, 300),
    na.value = "white"
  ) +
  theme_map() +
  labs(
    fill = expression("Monthly Average" ~ PM[2.5] ~ "(μg/m³)"),
    title = expression("Block-level Monthly Average" ~ PM[2.5] ~ "Pollution in India"),
    subtitle = as.character(ref_yr[i]),
    caption = expression("Graphic · Aarsh Batra | github.com/AarshBatra/biteSizedAQ")
  ) +
  theme(
    legend.position = "bottom",  # Position legend at bottom of plot
    legend.justification = "center",  # Center the legend
    legend.direction = "horizontal",
    legend.key.width = unit(3, "cm"),
    legend.box.margin = margin(0, 0, 15, 0),  # Add space below legend
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 13, face = "bold", family = "Mono"),
    plot.title = element_text(size = 20, family = "Mono", hjust = 0),
    plot.subtitle = element_text(size = 18, family = "Mono", hjust = 0, face = "bold", margin = margin(b = 1.5, unit = "cm")),
    legend.title = element_text(size = 14, family = "Mono", margin = margin(t = 0.2, unit = "cm")),
    legend.text = element_text(size = 12, family = "Mono"),
    plot.caption = element_text(size = 11, family = "Mono", margin = margin(t = 0.8, unit = "cm"), hjust = 0.5),
    plot.margin = margin(t = 0.3, r = 1, b = 0.3, l = 0.8, unit = "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~month, nrow = 3)
ggsave(paste0("./22.ind.block.pm2.5.sat.data.processing.1998.2024/monthly.india.pm2.5.snapshot.all.yrs/mth_avg_pm2.5_india_snapshot_", ref_yr[i], ".png"), plt1, width = 10, height = 10, dpi = 520)

print(paste0(i, " out of ", length(ref_yr), " year snapshot saved!"))
}


#### plot for selected years for all months for Delhi-----------------------

# Create your main plot 
plt1 <- final_all_years_cbind_df_sh_mapped_long %>%
  filter(year %in% c(1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016, 2018, 2020, 2022, 2024), state_name %in% c("nct of delhi")) %>%
  mutate(month = factor(month, levels = str_to_lower(month.name), labels = month.abb)) %>%
  left_join(
    shrug_master_subdist_shp %>%
      dplyr::select(state_name, district_name, subdistrict_name, geom), 
    by = c("state_name", "district_name", "subdistrict_name")
  ) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = avg_pm2.5), color = NA) +
  scale_fill_stepsn(
    colors = rev(viridisLite::magma(7)),
    breaks = c(0, 40, 80, 120, 200, 300),
    limits = c(0, 300),
    na.value = "white"
  ) +
  theme_map() +
  labs(
    fill = expression("Monthly Average" ~ PM[2.5] ~ "(μg/m³)"),
    title = expression("Block-level Monthly Average" ~ PM[2.5] ~ "Pollution in NCT of Delhi Blocks"),
    caption = expression("Graphic · Aarsh Batra | github.com/AarshBatra/biteSizedAQ")
  ) +
  theme(
    legend.position = "bottom",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.key.width = unit(3, "cm"),
    legend.box.margin = margin(0, 0, 15, 0),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, family = "Mono"),
    plot.title = element_text(size = 16, family = "Mono", hjust = 0, margin = margin(t = 0.5, b = 0.5, unit = "cm")),
    legend.title = element_text(size = 14, family = "Mono", margin = margin(t = 0.2, b = 0.1, unit = "cm")),
    legend.text = element_text(size = 12, family = "Mono"),
    plot.caption = element_text(size = 9, family = "Mono", margin = margin(t = 0.8, unit = "cm"), hjust = 0.5), 
    plot.margin = margin(l = 0.5, r = 0.5, unit = "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  facet_grid(year ~ month, switch = "y")   


# Save the map with high resolution
ggsave("./22.ind.block.pm2.5.sat.data.processing.1998.2024/experimental.plots/delhi_map_inset.png", plt1, width = 14, height = 14, dpi = 300)


```

## Testing Block [IGNORE]
```{r eval=FALSE, include=FALSE}

# Simple PM2.5 Monthly Stripes Visualization
# This script creates a simple climate stripe visualization for PM2.5 data
# showing 12 months for a specific block and year

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(viridis)  # For the magma color palette

# -------------------------------------------------------
# PART 1: SIMPLE FUNCTION FOR SINGLE BLOCK-YEAR STRIPES
# -------------------------------------------------------

create_monthly_stripes <- function(data, selected_state, selected_district, selected_block, selected_year) {
  # Filter data for the specific block and year
  filtered_data <- data %>% 
    filter(
      state_name == selected_state,
      district_name == selected_district,
      subdistrict_name == selected_block,
      year == selected_year
    ) %>%
    # Ensure months are in correct order
    arrange(month)
  
# Create the plot with custom tile width
  p <- ggplot(filtered_data, aes(x = factor(month), y = 1, fill = avg_pm2.5)) +
    # Set width of tiles (0.9 means 90% of available space)
    geom_tile(width = 0.9) +
    scale_fill_viridis_c(
      name = "PM2.5 (μg/m³)",
      option = "magma",      # Magma color palette
      direction = 1          # Direction: 1 for dark=high, -1 for dark=low
    ) +
    scale_x_discrete(
      breaks = 1:12,
      labels = month.abb     # Use abbreviated month names (Jan, Feb, etc.)
    ) +
    labs(
      title = paste("PM2.5 Monthly Pattern:", selected_block, selected_year),
      subtitle = paste("Population:", format(filtered_data$subdistrict_population[1], big.mark=",")),
      x = NULL, 
      y = NULL
    ) +
    theme_minimal() +
    theme(
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10, face = "bold")
    )
  
  return(p)
}

create_monthly_stripes(final_all_years_cbind_df_sh_mapped_long, selected_state =  "uttar pradesh", selected_district = "ghaziabad", selected_block = "modinagar", selected_year = "1998")

# -------------------------------------------------------
# PART 2: EXAMPLE USAGE WITH SAMPLE DATA
# -------------------------------------------------------

# Create sample data (replace this with your actual data)
set.seed(123)
sample_data <- expand.grid(
  state_name = "State1",
  district_name = "District1",
  subdistrict_name = paste0("Block", 1:5),
  year = 2020:2024,
  month = 1:12
)

sample_data$subdistrict_population <- round(runif(nrow(sample_data), 10000, 100000))

# Create PM2.5 data with seasonal pattern
sample_data$avg_pm2.5 <- 30 + 
                         # Year-over-year increase
                         (sample_data$year - 2020) * 5 + 
                         # Seasonal pattern (higher in winter months)
                         ifelse(sample_data$month %in% c(11, 12, 1, 2), 40, 
                           ifelse(sample_data$month %in% c(5, 6, 7, 8), -10, 0))

# Add some random variation
sample_data$avg_pm2.5 <- sample_data$avg_pm2.5 + rnorm(nrow(sample_data), mean = 0, sd = 10)

# Example: Create stripes for Block1 in 2022
example_plot <- create_monthly_stripes(sample_data, "Block1", 2022)
print(example_plot)

# Save plot (uncomment to use)
# ggsave("./22.ind.block.pm2.5.sat.data.processing.1998.2024/experimental.plots/Block1_2022_pm25_stripes.png", example_plot, width = 8, height = 2, dpi = 300)

# -------------------------------------------------------
# PART 3: SHINY APP SKELETON
# -------------------------------------------------------

# Uncomment and use this code as a starting point for your Shiny app

# library(shiny)
# 
# # UI
# ui <- fluidPage(
#   titlePanel("PM2.5 Monthly Climate Stripes"),
#   
#   sidebarLayout(
#     sidebarPanel(
#       # Block selection dropdown
#       selectInput("block", "Select Block:",
#                  choices = unique(sample_data$subdistrict_name)),
#       
#       # Year selection slider
#       sliderInput("year", "Select Year:",
#                  min = min(sample_data$year),
#                  max = max(sample_data$year),
#                  value = min(sample_data$year),
#                  step = 1,
#                  sep = ""),
#       
#       # Optional: Add filtering by state/district if needed
#       # selectInput("district", "Select District:",
#       #           choices = unique(sample_data$district_name))
#     ),
#     
#     mainPanel(
#       # Output plot
#       plotOutput("stripesPlot"),
#       
#       # Optional: Add summary statistics
#       verbatimTextOutput("summaryStats")
#     )
#   )
# )
# 
# # Server
# server <- function(input, output) {
#   # Reactive data filtering
#   filtered_data <- reactive({
#     sample_data %>%
#       filter(
#         subdistrict_name == input$block,
#         year == input$year
#       )
#   })
#   
#   # Generate the stripes plot
#   output$stripesPlot <- renderPlot({
#     create_monthly_stripes(sample_data, input$block, input$year)
#   })
#   
#   # Optional: Generate summary statistics
#   output$summaryStats <- renderPrint({
#     data_summary <- filtered_data() %>%
#       summarize(
#         avg = mean(avg_pm2.5, na.rm = TRUE),
#         min = min(avg_pm2.5, na.rm = TRUE),
#         max = max(avg_pm2.5, na.rm = TRUE)
#       )
#     
#     cat("Average PM2.5:", round(data_summary$avg, 1), "μg/m³\n")
#     cat("Range:", round(data_summary$min, 1), "to", round(data_summary$max, 1), "μg/m³\n")
#   })
# }
# 
# # Run the app
# shinyApp(ui = ui, server = server)

#### [experimetation] number of blocks exceeding x micrograms per cubic meter

# Set threshold
threshold <- 5

# Add column for number of months exceeding threshold
df1 <- final_all_years_cbind_df_sh_mapped_wide %>%
  rowwise() %>%
  mutate(
    months_exceeding_threshold = sum(c_across(starts_with("avg_pm2.5_")) > threshold, na.rm = TRUE)
  ) %>%
  ungroup()


# plt <- df1 %>%
#   filter(months_exceeding_threshold > 300)
#   ggplot() + 
#   geom_histogram(mapping = aes(x = months_exceeding_threshold), color = "white") 
  
  
# Create summary data for plotting
plot_data <- tibble(threshold = 0:max(df1$months_exceeding_threshold)) %>%
  rowwise() %>%
  mutate(
    percent_blocks_exceeding = mean(df1$months_exceeding_threshold > threshold) * 100
  ) %>%
  ungroup()

# Create the plot
plt <- ggplot(plot_data, aes(x = threshold, y = percent_blocks_exceeding)) +
  geom_line(color = "steelblue", size = 1) +
  labs(
    title = paste0("Share of Indian Blocks exceeding WHO annual average PM2.5 guideline"),
     subtitle = paste0("WHO annual average PM2.5 safe limit: ", threshold,  " μg/m³"),
    x = paste0("Number of months (1998 to 2024) exceeding 5 μg/m³"),
    y = "Percent of Blocks (%)", 
    caption = "*Note:There are a total of 324 months from 1998 to 2024"
  ) +
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0, margin = margin(t = 1, unit = "cm"), face = "italic", size = 9), 
        plot.subtitle = element_text(hjust = 0, margin = margin(b = 0.8, unit = "cm"), face = "italic", size = 10), 
             plot.title = element_text(hjust = 0, margin = margin(b = 0.2, unit = "cm"), face = "bold", size = 13), 
        axis.title.x = element_text(size = 11, margin = margin(t = 0.5, unit = "cm")), 
        axis.title.y = element_text(size = 11, margin = margin(r = 0.5, unit = "cm")), 
        axis.text = element_text(size = 10))



#### Delhi block level names map------------------------------

library(ggspatial)

# Extract the Delhi shapefile
delhi_sf <- shrug_master_subdist_shp %>%
  filter(state_name %in% c("nct of delhi"))

# Get the centroid coordinates for labeling
# This avoids font issues with geom_sf_text
delhi_centroids <- delhi_sf %>%
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame() %>%
  bind_cols(
    subdistrict_name = delhi_sf$subdistrict_name
    
  )

delhi_map <- ggplot() +
  # Base layer with clean outline and subtle background
  geom_sf(
    data = delhi_sf,
    fill = "gray98",
    color = "gray50",
    size = 0.3
  ) +
  # Add improved labels with better styling and placement
  geom_label_repel(
    data = delhi_centroids,
    aes(x = X, y = Y, label = subdistrict_name),
    color = "gray15",
    fill = alpha("white", 0.9),
    label.size = 0.15,
    size = 10,
    segment.color = "gray60",
    segment.size = 0.2,
    segment.alpha = 0.7,
    box.padding = 0.35,
    point.padding = 0.5,
    force = 10,
    max.overlaps = 40,
    min.segment.length = 0,
    label.padding = 0.15,
    label.r = unit(0.15, "lines"),
    seed = 42 # For reproducible placement
  ) +
  # Add subtle dots for centroids
  geom_point(
    data = delhi_centroids,
    aes(x = X, y = Y),
    size = 0.4,
    color = "gray40",
    alpha = 0.7
  ) +
  # Customize the coordinate appearance
  coord_sf(expand = FALSE) +
  # Add scale bar with improved styling
  annotation_scale(
    location = "br",
    width_hint = 0.15,    # Keep original width
    style = "bar",        # Bar style for better visibility
    pad_x = unit(0.4, "cm"),
    pad_y = unit(0.4, "cm"),
    text_family = "franklin",
    text_col = "black",   # Darker text
    line_col = "black",   # Darker line
    bar_cols = c("black", "white"), # High contrast colors
    height = unit(0.25, "cm"), # Slightly taller
    text_cex = 1.1        # Slightly larger text
  ) +
  # Add north arrow with improved styling
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    pad_x = unit(0.4, "cm"),
    pad_y = unit(0.4, "cm"),
    style = north_arrow_fancy_orienteering(
      line_col = "black",
      fill = "gray30",
      text_col = "black",
      text_size = 14,
      line_width = 1.5
    ),
    height = unit(1.5, "cm"),
    width = unit(1.5, "cm")
  ) +
  # Clean, minimalist theme
  theme_map() +
  labs(title = "NCT of Delhi Blocks") +
  theme(
    plot.title = element_text(size = 46, hjust = 0.5, margin = margin(b = 1, t = 0.5, unit = "cm")), 
    panel.border = element_rect(color = "black", fill = NA, size = 0.3), 
    plot.background = element_rect(fill = "white", color = "white")
  )


#### Sanity Check: state, year wise population weight annual average pm2.5 (for comaprison with ACAG estimates)
final_all_years_cbind_df_sh_mapped_long %>% group_by(state_name, year) %>% mutate(st_pop = sum(subdistrict_population, na.rm = TRUE), pop_wt = subdistrict_population/st_pop, pop_wt_pm2.5 = pop_wt*avg_pm2.5) %>% summarise(avg_st_yr_pm2.5 = sum(pop_wt_pm2.5, na.rm = TRUE)) %>% ungroup() %>% View()
  

```

